\chapter{Conclusion and future work}
\label{chapter:conclusion}

My primary object was to create the Diplomatiq social network software system. To build and publish the system, this involved creating several other supportive libraries, and a company-level, production-grade server infrastructure as well, with several kinds of supportive infrastructure.

The initial version of the Diplomatiq software was released as a publicly available web application, and is suitable for organizing MUN conferences. The application was implemented upon a modern, layered security architecture, which provides cryptographic assurances in terms of application and data security.

\section{Summary of contributions}

I have achieved the following engineering contributions:

\begin{itemize}
\item I designed, built, secured and paid a company-level production infrastructure for the development, testing, production operation and maintenance of Diplomatiq, including several kinds of supportive infrastructure.
\item I designed an application security framework addressing authentication, authorization, and data protection, which provides cryptographic assurances regarding access control and the confidentiality and integrity of sensitive data.
\item I designed, implemented and published an elementary version of the Diplomatiq social network software system as a client-server application, using graph database technologies and the aforementioned security framework.
\item I developed several supportive libraries outside the Diplomatiq software along the way. I published the built artefacts of these libraries with detailed documentation, for free use in the open-source community.
\item I published the source code of all my contributions as separate open-source projects, centralized under one project organization, called Diplomatiq.
\end{itemize}

\section{Future work}

The goal of the work described in this thesis is to build a solid foundation for Diplomatiq, the social network software system, and also Diplomatiq, the prospective company, developing and maintaining the application in the future. The foundation was created along with a minimal feature set. However, much more features need to be implemented into Diplomatiq, so it can be a truly powerful tool for organizing Model United Nations conferences, and for enabling delegates to prepare to conferences better.
